Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-02-09 paper.tex
no more errors

  \node at (nn) {
    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.15,shift={+(-1,2.5)}}]
      \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
      \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
      \node[rectangle,width=2cm] at (1,1) {};
      \foreach \name / \y in {1,...,4}
          \node[neuron] (I-\name) at (0,-\y) {};
      \foreach \name / \y in {1,...,3}
          \node[neuron] (H-\name) at (1,-\y-0.5) {};
      \foreach \name / \y in {1,...,4}
          \node[neuron] (O-\name) at (2,-\y) {};
      \foreach \source in {1,...,4}
          \foreach \dest in {1,...,3}
              \draw [-latex] (I-\source) -- (H-\dest);
      \foreach \source in {1,...,3}
          \foreach \dest in {1,...,4}
              \draw [-latex] (H-\source) -- (O-\dest);
    \end{tikzpicture}
  };
  \node[shift={+(0,-0.55)}] at (nn) { $q(x)$ };

@@

%% \begin{figure}
%% \centering
%% \begin{tikzpicture}
%%   \begin{scope}[shift={(2.5,0)}]
%%     \node[obs] at (3,3) (dx){$\mathcal{D}$};
%%     \node[latent] at (3.5,1.75) (zp){$p$};
%%     \node[obs] at (4,3) (tx){$\theta^{(x)}$};
%%     \node[obs] at (3.5,0.7) (xp) {$x$};
%%     \node[align = center] at ([yshift = -0.6cm,xshift = 0.5cm]xp.south) {Wake: Infer $p$};
%%     \draw [->,red] (xp.east) to[out = 0,in = 0] node(nn){} (tx.east);
%%     \draw [->,red] (tx) -- (zp);
%%     \draw [->] (dx) -- (zp);
%%     \node at (nn) {
%%       \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
%%         \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
%%         \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
%%         \node[rectangle,width=2cm] at (1,1) {};
%%         \foreach \name / \y in {1,...,4}
%%             \node[neuron] (I-\name) at (0,-\y) {};
%%         \foreach \name / \y in {1,...,3}
%%             \node[neuron] (H-\name) at (1,-\y-0.5) {};
%%         \foreach \name / \y in {1,...,4}
%%             \node[neuron] (O-\name) at (2,-\y) {};
%%         \foreach \source in {1,...,4}
%%             \foreach \dest in {1,...,3}
%%                 \draw [-latex] (I-\source) -- (H-\dest);
%%         \foreach \source in {1,...,3}
%%             \foreach \dest in {1,...,4}
%%                 \draw [-latex] (H-\source) -- (O-\dest);
%%       \end{tikzpicture}
%%     };
%%     \node[shift={+(0,-0.65)}] at (nn) [fill=white]{ $q(x)$ };
%%   \end{scope}

%%   \begin{scope}[shift={(0.7,-4.2)}]
%%     \node[obs] at (3,3) (dt){$\mathcal{D},\theta$};
%%     \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]dt.south) (p2){$p_2$};
%%     \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p2.south) (x2){$x_2$};
%%     \draw [->] (p2.south) -- (x2.north);
%%     \draw [->] (dt.south) -- (p2.north);

%%     \node[obs] at ([yshift = 0cm,xshift = -0.5cm]p2.west) (p1){$p_1$};
%%     \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p1.south) (x1){$x_1$};
%%     \draw [->] (p1.south) -- (x1.north);
%%     \draw [->] (dt.south) -- (p1.north);

%%     \node[obs] at ([yshift = 0cm,xshift = 0.5cm]p2.east) (p3){$p_3$};
%%     \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p3.south) (x3){$x_3$};
%%     \draw [->] (p3.south) -- (x3.north);
%%     \draw [->] (dt.south) -- (p3.north);

%%     \node at ([yshift = 0cm,xshift = 0.3cm]p3.east) {$\cdots $};
%%     \node at ([yshift = 0cm,xshift = 0.3cm]x3.east) {$\cdots $};

%%     \node[obs] at ([yshift = 0cm,xshift = 1.3cm]p3.east) (zp){$p$};
%%     \node[obs] at ([yshift = 0cm,xshift = 1.3cm]x3.east) (xp) {$x$};
%%     \draw [->] (zp.south) -- (xp.north);
%%     \plate {}{(zp)(xp)}{$x\in X$};

%%     \draw[dashed,cyan,very thick] ([yshift = 0.5cm,xshift = -0.5]p1.west) rectangle  ([yshift = -0.5cm,xshift = -0.5]xp.east);
%%     \node[align = center] at ([yshift = -1cm,xshift = 0.6cm]x2.south) {Sleep-R: Train $q$\\training data: cyan $(x,p)$};
%%   \end{scope}

%%   \begin{scope}[shift={(7.7,-4)}]
%%     \node[latent] at (0.5,3) (d){$\mathcal{D}$};
%%     \node[latent] at (1.5,3) (t){$\theta$};
%%     \node[obs] at (1,1.75) (z){$p$};
%%     % \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
%%     \node[obs] at (1,0.5) (x) {$x$};
%%     \edge {z}{x};
%%     \edge {d,t}{z};
%%     \plate {}{(z)(x)}{$x\in X$};
%%     \node[align = center] at ([yshift = -1cm]x.south) {Sleep-G: Induce $(\mathcal{D},\theta})$};
%%   \end{scope}

%% \end{tikzpicture}
%% \end{figure}

@@

\begin{algorithm}[tb]
   \caption{Generative model over programs}
   \label{programGenerativeModel}
   \begin{algorithmic}
     \STATE \textbf{function} sampleProgramFromDSL$(\mathcal{D}, \theta, \tau)$:
  \STATE {\bfseries Input:} DSL $\mathcal{D}$, weight vector $\theta$, type $\tau$
  \STATE \textbf{Output:} a program whose type unifies with $\tau$
  \STATE \textbf{return} sample$(\mathcal{D}, \theta, \varnothing, \tau)$
\STATE
     \STATE \textbf{function} sample$(\mathcal{D}, \theta, \mathcal{E}, \tau)$:
  \STATE {\bfseries Input:} DSL $\mathcal{D}$, weight vector $\theta$, environment $\mathcal{E}$, type $\tau$
  \STATE \textbf{Output:} a program whose type unifies with $\tau$
  \IF{$\tau = \alpha\to\beta$}
  \STATE var $\gets$ an unused variable name
  \STATE body $\sim$ sample$(\mathcal{D},\theta,\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$
   \STATE \textbf{return} \code{(lambda (}var\code{) }body\code{)}
   \ENDIF
   %   \ELSE
   \STATE $\text{primitives} \gets\{p | &p: \tau' \in \mathcal{D}\cup\mathcal{E}$
   \STATE \hspace{2.5cm}$\text{if }\tau\text{ can unify with yield}(\tau') \} $

   \STATE Draw $e\sim \text{primitives}$, w.p. $\propto\theta_e$ if $e\in \mathcal{D}$
   \STATE \hspace{3.1cm}w.p. $\propto\frac{\theta_{var}}{|\text{variables}|}$ if $e\in \mathcal{E}$
   \STATE Unify $\tau$ with yield$(\tau')$.
   \STATE $\left\{\alpha_k \right\}_{k = 1}^K\gets\text{argTypes}(\tau')$ 
%   \STATE unify$(\tau,\beta)$
   \FOR{$k=1$ {\bfseries to} $K$}
 \STATE $a_k\sim\text{sample}(\mathcal{D},\theta,\mathcal{E},\alpha_k)$
 \ENDFOR
 \STATE \textbf{return} \code{(}$e\;a_1\; a_2\; \cdots\; a_K$\code{)}
 \STATE\textbf{where:}
 \STATE  yield$(\tau) = \begin{cases}
\text{yield}(\beta)   &\text{ if }$\tau = \alpha\to \beta$\\
\tau   &\text{ otherwise.}
 \end{cases}$ 
 \STATE\hspace{0cm} argTypes$(\tau) = \begin{cases}
[\alpha] + \text{argTypes}(\beta)   &\text{ if }$\tau = \alpha\to \beta$\\
[]   &\text{ otherwise.}
 \end{cases}$
\end{algorithmic}
\end{algorithm}

@@

\begin{figure}\centering
  \begin{circuitikz} \draw
    (0,0) node[port] (x) {$x$}
    (x)++(0,-0.3) node[port] (y) {$y$}
    (y)++(0,-0.5) node[port] (z) {$z$}

    (x)++(1.7,-0.15) node[nand port, scale = 0.5] (xy) {}
    (xy) ++ (0,-1) node[nand port, scale = 0.5] (xz) {}
    (xz) ++ (1,0.2) node[nand port, scale = 0.5] (zxz) {}
    (zxz) ++ (1,0.2) node[nand port, scale = 0.5] (m) {}
    (x.out) -- (xy.in 1)
    (y.out) -- (xy.in 2)

    (x.out) -| (xz.in 1)
    (z.out) -- (xz.in 2)

    (xz.out) -- (zxz.in 2)
    (z.out) |- (zxz.in 1)

@@

    (m)++(1.2,0) node [port] (l) {\textsc{MUX}$(x,y,z)$}
    ;
%% (0,0) node[and port] (myand2) {}
%% (2,1) node[xnor port] (myxnor) {}
%% (myand1.out) -- (myxnor.in 1)
%% (myand2.out) -- (myxnor.in 2);
  \end{circuitikz}
  \begin{circuitikz} \draw
    (0,0) node[port] (a) {$a$}
    (a)++(0,-0.7) node[port] (b) {$b$}

@@

    (a.out) -| (ab1.in 1)
    (a.out) -| (ab1.in 2)
    (b.out) -| (ab2.in 1)
    (b.out) -| (ab2.in 2)

@@

    (o)++(0.8,0) node [port] (l) {\textsc{OR}$(a,b)$}
    ;
  \end{circuitikz}
  \begin{circuitikz}     \node(m)[draw] at (1,-1) {MUX}; \draw
    (0,0) node[port] (w) {$w$}
    (w)++(0,-0.5) node[port] (x) {$x$}
    (x)++(0,-0.5) node[port] (y) {$y$}
    (y)++(0,-0.5) node[port] (z) {$z$}

    (z.out) -| (m.south)
    (y.out) -| (m.west)
    (x.out) -| (m.north)

>>>>

  \node at (nn) {
    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.15,shift={+(-1,2.5)}}]
      \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
      \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
      \node[rectangle] at (1,1) {};
      \foreach \name / \y in {1,...,4}
          \node[neuron] (I-\name) at (0,-\y) {};
      \foreach \name / \y in {1,...,3}
          \node[neuron] (H-\name) at (1,-\y-0.5) {};
      \foreach \name / \y in {1,...,4}
          \node[neuron] (O-\name) at (2,-\y) {};
      \foreach \source in {1,...,4}
          \foreach \dest in {1,...,3}
              \draw [-latex] (I-\source) -- (H-\dest);
      \foreach \source in {1,...,3}
          \foreach \dest in {1,...,4}
              \draw [-latex] (H-\source) -- (O-\dest);
    \end{tikzpicture}
  };
  \node[shift={+(0,-0.55)}] at (nn) { $q(x)$ };

@@

@@

\begin{algorithm}[tb]
   \caption{Generative model over programs}
   \label{programGenerativeModel}
   \begin{algorithmic}
     \STATE \textbf{function} sampleProgramFromDSL$(\mathcal{D}, \theta, \tau)$:
  \STATE {\bfseries Input:} DSL $\mathcal{D}$, weight vector $\theta$, type $\tau$
  \STATE \textbf{Output:} a program whose type unifies with $\tau$
  \STATE \textbf{return} sample$(\mathcal{D}, \theta, \varnothing, \tau)$
\STATE
     \STATE \textbf{function} sample$(\mathcal{D}, \theta, \mathcal{E}, \tau)$:
  \STATE {\bfseries Input:} DSL $\mathcal{D}$, weight vector $\theta$, environment $\mathcal{E}$, type $\tau$
  \STATE \textbf{Output:} a program whose type unifies with $\tau$
  \IF{$\tau = \alpha\to\beta$}
  \STATE var $\gets$ an unused variable name
  \STATE body $\sim$ sample$(\mathcal{D},\theta,\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$
   \STATE \textbf{return} \code{(lambda (}var\code{) }body\code{)}
   \ENDIF
   %   \ELSE
   \STATE $\text{primitives} \gets\{p | p: \tau' \in \mathcal{D}\cup\mathcal{E}$ \\
     \hspace*{6.9em}$\text{if }\tau\text{ can unify with yield}(\tau') \} $

   \STATE Draw $e\sim \text{primitives}$, w.p. $\propto\theta_e$ if $e\in \mathcal{D}$
   \STATE \hspace*{8.8em}w.p. $\propto\frac{\theta_{var}}{|\text{variables}|}$ if $e\in \mathcal{E}$
   \STATE Unify $\tau$ with yield$(\tau')$.
   \STATE $\left\{\alpha_k \right\}_{k = 1}^K\gets\text{argTypes}(\tau')$ 
%   \STATE unify$(\tau,\beta)$
   \FOR{$k=1$ {\bfseries to} $K$}
 \STATE $a_k\sim\text{sample}(\mathcal{D},\theta,\mathcal{E},\alpha_k)$
 \ENDFOR
 \STATE \textbf{return} \code{(}$e\;a_1\; a_2\; \cdots\; a_K$\code{)}
 \STATE\textbf{where:}
 \STATE \quad yield$(\tau) = \begin{cases}
    \text{yield}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
    \tau   &\text{ otherwise.}
 \end{cases}$ 
 \STATE \quad argTypes$(\tau) = \begin{cases}
    [\alpha] + \text{argTypes}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
    []   &\text{ otherwise.}
 \end{cases}$
\end{algorithmic}
\end{algorithm}

@@

\begin{figure}\centering
  \begin{circuitikz} \draw
    (0,0) node (x) {$x$}
    (x)++(0,-0.3) node (y) {$y$}
    (y)++(0,-0.5) node (z) {$z$}

    (x)++(1.7,-0.14) node[nand port, scale = 0.5] (xy) {}
    (xy) ++ (0,-1) node[nand port, scale = 0.5] (xz) {}
    (xz) ++ (1,0.2) node[nand port, scale = 0.5] (zxz) {}
    (zxz) ++ (1,0.2) node[nand port, scale = 0.5] (m) {}
    (x) -- (xy.in 1)
    (y) -- (xy.in 2)

    (x) -| (xz.in 1)
    (z.east) -- (xz.in 2)

    (xz.out) -- (zxz.in 2)
    (z.east) |- (zxz.in 1)

@@

    (m)++(1.2,0) node (l) {\textsc{MUX}$(x,y,z)$}
    ;
%% (0,0) node[and port] (myand2) {}
%% (2,1) node[xnor port] (myxnor) {}
%% (myand1.out) -- (myxnor.in 1)
%% (myand2.out) -- (myxnor.in 2);
  \end{circuitikz}
  \begin{circuitikz} \draw
    (0,0) node (a) {$a$}
    (a)++(0,-0.7) node (b) {$b$}

@@

    (a.east) -| (ab1.in 1)
    (a.east) -| (ab1.in 2)
    (b.east) -| (ab2.in 1)
    (b.east) -| (ab2.in 2)

@@

    (o)++(0.8,0) node (l) {\textsc{OR}$(a,b)$}
    ;
  \end{circuitikz}
  \begin{circuitikz}     \node(m)[draw] at (1,-1) {MUX}; \draw
    (0,0) node (w) {$w$}
    (w)++(0,-0.5) node (x) {$x$}
    (x)++(0,-0.5) node (y) {$y$}
    (y)++(0,-0.5) node (z) {$z$}

    (z.east) -| (m.south)
    (y.east) -| (m.west)
    (x.east) -| (m.north)
